Errorless irrationality: removing error-driven components from the inverse base-rate effect paradigm

### 2023-01-03 docs/manuscript.tex
wrap up introduction and move on to experiment 1

In this work, we intend to test this basic assumption of models of the IBRE.
In the following two experiments, we will gradually remove components from the design traditionally associated with prediction error.
In our first attempt, building on the observational learning condition of Experiment 2 in \citeA{johansen2007paradoxical}, we implement the canonical IBRE design with a caveat that category labels are presented in unison with features.

@@

To our knowledge, there is only one attempt to implement the IBRE procedure without explicit feedback.
\citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational learning condition, where category labels were presented with features at the same time to participants.
Their design involves disjoint-cues (where categories shared no features in common), while the canonical design depends on a shared feature during training that facilitates attentional relocation.
This attentional tuning in turn pushes responding towards the rare label.
Their design was optimised to investigate the hypothesised assymetric cognitive representation of the two categories - one of the assumptions.
As a result, the only instance when they observed a rare bias was when common features presented in compound during training were paired with a rare feature presented by itself during training.
This provided evidence for the assymetric cognitive represetation hypothesised to develop during training.
In order to investigate the role of prediction error in response to feedback, we need to tweak their observational learning condition to conform to a more canonical implementation of the procedure.
A simple neural-network explanation for the rare bias \citeA{johansen2007paradoxical} observed is that the rare feature developed stronger connections with the category label.
Compound features share the connection with the category label, so any update to these connections dissipate between them.
There is no need to relocate attention to reduce errors, therefore attention will not bias responding towards the rare label.

The only two studies directly looking at error-driven processes in the IBRE are \citeA{inkster2022neural} and \citeA{wills2014attention}.
\citeA{inkster2022neural} carried out a direct investigation into brain regions underlying error-driven learning in the IBRE.
They observed

@@

Below, we detail our first attempt to test whether we could observe the rare response bias without an explicit error-driven psychological mechanism.
The design component which is most likely to result in any error-driven tuning is feedback.
To remove feedback, we will present category labels with their respective features.

@@

@@

@@

\begin{table}[H]
  \begin{center}
    \caption{Stuff. \\}
    \label{tab:results-exp1}
    \vskip 0.12in
    \begin{tabular}{rcc}
      \hline
      & $P(common)$ & $P(rare)$ \\
      \hline
      A & 0.69 & 0.31\\
      AB & 0.93 & 0.07\\
      AC & 0.10 & 0.90\\
      B & 0.93 & 0.07\\
      \textbf{BC} & \textbf{0.33} & \textbf{0.66}\\
      C & 0.06 & 0.94\\
    \end{tabular}
  \end{center}
\end{table}

$M = 0.66$, 95\% HDI $[0.62, 0.71]$, $\mathrm{BF}_{10} = 6.63 \times 10^{7}$

@@

@@

\begin{table}[H]
  \begin{center}
    \caption{Caption.\\}
    \label{tab:results-exp2}
    \vskip 0.12in
    \begin{tabular}{rcc}
      \hline
       & $P(common)$ & $P(rare)$ \\
      \hline
      A & 0.76 & 0.24\\
      AB & 0.93 & 0.07\\
      AC & 0.11 & 0.89\\
      B & 0.92 & 0.08\\
      \textbf{BC} & \textbf{0.34} & \textbf{0.65}\\
      C & 0.09 & 0.91\\
    \end{tabular}
  \end{center}
\end{table}

$M = 0.65$, 95\% HDI $[0.55, 0.76]$, $\mathrm{BF}_{10} = 8.13$

>>>>

In this work, we intend to test this basic assumption of models of the IBRE.
In the following two experiments, we will gradually remove components from the design traditionally associated with prediction error.
Our overarching goal is to investigate whether we can still observe the IBRE, even if we experimentally remove a crucial assumption of already existing accounts.
In our first attempt, building on the observational learning condition of Experiment 2 in \citeA{johansen2007paradoxical}, we implement the canonical IBRE design with a caveat that category labels are presented in unison with features.

@@

To our knowledge, there is only one attempt to implement the IBRE procedure without explicit feedback.
\citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational learning condition, where category labels were presented with features at the same time to participants.
Their design involved disjoint-cues (where categories shared no features in common), while the canonical design depends on a shared feature during training that facilitates attentional relocation.
This attentional tuning in turn pushes responding towards the rare label.
Their design was optimised to investigate the hypothesised assymetric cognitive representation of the two categories - one of the assumptions.
As a result, the only instance when they observed a rare bias was when common features presented in compound during training were paired with a rare feature presented by itself during training.
This provided evidence for the assymetric cognitive represetation hypothesised to develop during training.
In order to investigate the role of prediction error in response to feedback, we need to tweak their observational learning condition to conform to a more canonical implementation of the procedure.
A simple neural-network explanation can account for this by positing that the rare feature developed stronger connections with the category label.
Compound features share the connection with the category label, so any update to these connections dissipate between them.
There is no need to relocate attention to reduce errors, therefore attention will not bias responding towards the rare label.
But stronger $C \to rare$ feature-label connections will result in a rare bias on BC compunds during test.
Our current study modified the canonical IBRE procedure in an attempt to remove error-driven components from the design - which includes a shared cue.
This contrasts \citeA{johansen2007paradoxical}, who tried to remove the shared cue that pushed attention towards C during training and resulted in the assymetric representation.
Here, we strictly focus on prediction error in relation to behaviour.
% I will also need to include Experiment 3 listed shared cue condition

% I need to go through some studies because I think there are at least one more, but I 
The only two studies directly looking at error-driven processes in the IBRE are \citeA{inkster2022neural} and \citeA{wills2014attention}.
\citeA{wills2014attention} observed posterior selection negativity and concurrent frontal positivity for C relative to B, which gave evidence for an error-driven selective attentional learning process.
\citeA{inkster2022neural} carried out a direct investigation into brain regions underlying error-driven learning in the IBRE.
Their ROI analysis explicitly targeted areas that were hypothesised to be involved in the computation of prediction error.
They showed that these areas exhibited greater activation during the test phase for C relative to B with a presence of a shared cue during training.
Both \citeA{wills2014attention} and \citeA{inkster2022neural} gave strong evidence in support of error-driven attentional learning accounts of the IBRE.
In our study, we will look for the effect while trying to take out prediction error from the experimental design.
In contrast, they looked for the neural substrates of error-driven accounts of the effect.

@@

Below, we detail our first attempt to test whether we could observe the rare response bias without an explicit error-driven psychological mechanism.
The design component which is most likely to result in any error-driven tuning is feedback.
To remove feedback, we will present category labels with their respective features.
We retain the sequential property of the experiment, which means that participants learn about feature and category relationships on a trial by trial basis.
% work out how models might predict rational responding

@@

\subsubsection*{Participants}

Recruited 169 participants.

\subsubsection*{Stimuli}

\subsubsection*{Procedure}

@@

\subsubsection*{Exclusion}

To match performance with

\subsubsection*{Analysis}

@@

125 ppt made it into the analysis.

\begin{table}[H]
  \begin{center}
    \caption{Stuff. \\}
    \label{tab:results-exp1}
    \vskip 0.12in
    \begin{tabular}{rcc}
      \hline
      & $P(common)$ & $P(rare)$ \\
      \hline
      A & 0.69 & 0.31 \\
      AB & 0.94 & 0.06 \\
      AC & 0.08 & 0.92 \\
      B & 0.94 & 0.06 \\
      \textbf{BC} & \textbf{0.33} & \textbf{0.67} \\
      C & 0.04 & 0.96 \\
    \end{tabular}
  \end{center}
\end{table}

$M = 0.67$, 95\% HDI $[0.62, 0.72]$, $\mathrm{BF}_{10} = 1.11 \times 10^{7}$

@@

\subsubsection*{Participants}

Recruited 65 participants.

\subsubsection*{Stimuli}

\subsubsection*{Procedure}

@@

30 made it into the analysis

\begin{table}[H]
  \begin{center}
    \caption{Caption.\\}
    \label{tab:results-exp2}
    \vskip 0.12in
    \begin{tabular}{rcc}
      \hline
       & $P(common)$ & $P(rare)$ \\
      \hline
      A & 0.78 & 0.22  \\
      AB & 0.95 & 0.05 \\
      AC & 0.09 & 0.91 \\
      B & 0.92 & 0.07  \\
      \textbf{BC} & \textbf{0.35} & \textbf{0.65} \\
      C & 0.08 & 0.92 \\
    \end{tabular}
  \end{center}
\end{table}

$M = 0.64$, 95\% HDI $[0.53, 0.75]$, $\mathrm{BF}_{10} = 4.09$
